---
title: "Data Wrangling"
author: "Alice Jiang"
date: "3/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
```

## Data Wrangling
```{r import_data}
bankruptcy_financials <- read.csv("data/bankruptcy.csv")
stock_price <- read.csv("data/stock price.csv")
stock_price <- stock_price[, -(2:4)]
oil <- read.csv("data/Brent Crude Oil.csv") # https://fred.stlouisfed.org/series/POILBREUSDM
cpi <- read.csv("data/Core CPI.csv") # https://fred.stlouisfed.org/series/CPILFESL
yield <- read.csv("data/3-month yield.csv") # https://fred.stlouisfed.org/series/DGS3MO#0
nonfarm_payroll <- read.csv("data/nonfarm payroll.csv")
```

```{r}
# dim(bankruptcy_financials)
# dim(stock_price)
```

```{r financial_data}
data <- left_join(bankruptcy_financials, stock_price, by = "Symbol")
# dim(data)
```

```{r year}
data <- data %>%
  mutate(
    Date..RT.Last. = strptime(data$Date..RT.Last., "%m/%d/%Y"),
    Year = year(Date..RT.Last.), # year of last trade
    Month = month(Date..RT.Last.), # month of last trade
  )
```

```{r subset_data}
# 2010-2021
data <- data %>%
  filter(Year >= 2010)
# dim(data)
```

```{r bankrupt}
data <- data %>%
  mutate(
    Bankrupt = ifelse(Bankruptcy.Status == "BANKRUPT", 1, 0) # bankruptcy indicator
  )
```

```{r convert_NA}
data[6:8] <- lapply(data[6:8], function(x) ifelse(x == "-", NA, as.character(x)))
data[11:216] <- lapply(data[11:216], function(x) ifelse(x == "-", NA, as.character(x)))
data[217:277] <- lapply(data[217:277], function(x) ifelse(x == "-", NA, as.character(x)))
```

```{r match1}
# matches the latest financial data available to each company based on their last trading year
# E.g. Total.Assets (2021) | Total.Assets.11 (2010)
match1 <- function(data, year, var) {
  # data, data$Year, "", 
  n <- rep(0, length(year))
  var_new <- rep("", length(year))
  output <- rep("", length(year))
  for (i in 1:length(year)) {
    while (TRUE) {
      n[i] <- 2021 - year[i] 
      #print(n[i])
      if (n[i] == 0) {
        var_new[i] <- var
      }
      else if (n[i] > 0) {
        var_new[i] <- paste0(var, ".", n[i])
      }
      #print(var_new[i])
      if (!is.na(data[[var_new[i]]][i])) { # data is available
        output[i] <- data[[var_new[i]]][i]
        break
      }
      if (n[i] == 11) { # loops to 2010 data 
        output[i] <- NA 
        break
      }
      year[i] <- year[i] - 1 # goes back one year if current data is NA
    }
  }
  return(output)
}
```

```{r financial_ratios}
# cols <- c("Total_Assets", "Net_Sales", "EV_to_Sales", "EBITDA", "ROA", "ROE", "Working_Capital", 
#           "Retained_Earnings", "Profit_Margin", "Debt_to_Assets", "Debt_to_Equity", "Interest_Coverage_Ratio",
#           "Current_Ratio", "Cash_From_Operations", "Total_Debt", "Debt_to_EBITDA", "Interest_Expense", "Market_Value")
# 
# names <- c("Total.Assets", "Net.Sales", "Entrpr.Value..Sales", "EBITDA.Oper.Income",
#            "Return.on.Avg.Total.Assets", "Return.on.Avg.Total.Equity", "Working.Capital.Total", 
#           "Retain.Earns..Accum.Deficit", "Net.Income.Margin", "Total.Debt..Total.Assets",
#           "Total.Debt..Equity", "Pretax.Int.Coverag",
#           "Current.Ratio", "Net.Cash.Flow..Oper", "Total.L.T...S.T.Debt", "Total.Debt..EBITDA", 
#           "Int.on.Debt.Expense", "Market.Value")

data <- data %>%
  mutate(Total_Assets = match1(data, data$Year, "Total.Assets"),
         Net_Sales = match1(data, data$Year, "Net.Sales"),
         EV_to_Sales = match1(data, data$Year, "Entrpr.Value..Sales"),
         EBITDA = match1(data, data$Year, "EBITDA.Oper.Income"),
         ROA = match1(data, data$Year, "Return.on.Avg.Total.Assets"), 
         ROE = match1(data, data$Year, "Return.on.Avg.Total.Equity"),
         Working_Capital = match1(data, data$Year, "Working.Capital.Total"),
         Retained_Earnings = match1(data, data$Year, "Retain.Earns..Accum.Deficit"),
         Profit_Margin = match1(data, data$Year, "Net.Income.Margin"),
         Debt_to_Assets = match1(data, data$Year, "Total.Debt..Total.Assets"),
         Debt_to_Equity = match1(data, data$Year, "Total.Debt..Equity"),
         Interest_Coverage_Ratio = match1(data, data$Year, "Pretax.Int.Coverag"),
         Current_Ratio = match1(data, data$Year, "Current.Ratio"),
         Cash_From_Operations = match1(data, data$Year, "Net.Cash.Flow..Oper"),
         Total_Debt = match1(data, data$Year, "Total.L.T...S.T.Debt"),
         Debt_to_EBITDA = match1(data, data$Year, "Total.Debt..EBITDA"),
         Interest_Expense = match1(data, data$Year, "Int.on.Debt.Expense"),
         Market_Value = match1(data, data$Year, "Market.Value")
  )
```

```{r match2}
# matches the latest stock price available to each company based on their last trading year
# E.g. Closing.Price_Dec (2020) | Closing.Price_Dec.11 (2009) 
# E.g. Closing.Price_Mar (2020) | Closing.Price_Mar.10 (2010) 
match2 <- function(data, year, var, start, end) {
  # data, data$Year, "", (2020), (2009)
  n <- rep(0, length(year))
  var_new <- rep("", length(year))
  output <- rep("", length(year))
  for (i in 1:length(year)) {
    if(year[i] == 2021) {
      year[i] <- 2020 # use 2020 data instead
    }
    while (TRUE) {
      n[i] <- start - year[i] # e.g. 2020
      #print(n[i])
      if (n[i] == 0) {
        var_new[i] <- var
      }
      else if (n[i] > 0) {
        var_new[i] <- paste0(var, ".", n[i])
      }
      #print(var_new[i])
      if (!is.na(data[[var_new[i]]][i])) { # data is available
        output[i] <- data[[var_new[i]]][i]
        break
      }
      if (n[i] == start - end) { # loops to 2009/2010 data 
        output[i] <- NA 
        break
      }
      year[i] <- year[i] - 1 # goes back one year if current data is NA
    }
  }
  return(output)
}
```


```{r stock_price}
# cols <- c("Closing_Price_Dec", "Closing_Price_Mar", "Closing_Price_Jun", "Closing_Price_Sep")
# 
# names <- c("Closing.Price_Dec", "Closing.Price_Mar", "Closing.Price_Jun", "Closing.Price_Sep")

data <- data %>%
  mutate(Closing_Price_Dec = match2(data, data$Year, "Closing.Price_Dec", 2020, 2009),
         Closing_Price_Mar = match2(data, data$Year, "Closing.Price_Mar", 2020, 2010),
         Closing_Price_Jun = match2(data, data$Year, "Closing.Price_Jun", 2020, 2010),
         Closing_Price_Sep = match2(data, data$Year, "Closing.Price_Sep", 2020, 2010)
  )
```

```{r match3}
# match year to latest available data based on revenue
year <- data$Year
n <- rep(0, length(year))
var_new <- rep("", length(year))
latest_year <- rep(0, length(year))
var <- "Net.Sales"
for (i in 1:length(year)) {
  while (TRUE) {
    n[i] <- 2021 - year[i]
    if (n[i] == 0) {
      var_new[i] <- var
    }
    else if (n[i] > 0) {
      var_new[i] <- paste0(var, ".", n[i])
    }
    if (!is.na(data[[var_new[i]]][i])) {
      latest_year[i] <- year[i] # year with the latest available data 
      break
    }
    if (n[i] == 11) {
      latest_year[i] <- NA # loops to 2010 data
      break
    }
    year[i] <-
      year[i] - 1 # goes back one year if current data is NA
  }
}
data$Latest_Year <- latest_year
```

## Select Universe of Companies (bankrupt and non-bankrupt)
```{r filter_data}
final_data <- data %>%
  select(Symbol, Name, Bankrupt, FactSet.Econ.Sector, FactSet.Ind, Entity.NAICS.Sector.Full, Year, Latest_Year, Month, 281:302, X3y.Price.Volatility, X5y.Price.Volatility)
# dim(final_data)

cols <- c("Total_Assets", "Net_Sales", "EV_to_Sales", "EBITDA", "ROA", "ROE", "Working_Capital",
          "Retained_Earnings", "Profit_Margin", "Debt_to_Assets", "Debt_to_Equity", "Interest_Coverage_Ratio",
          "Current_Ratio", "Cash_From_Operations", "Total_Debt", "Debt_to_EBITDA", "Interest_Expense")

final_data <- final_data %>%
  filter_at(vars(cols), any_vars(!is.na(.))) # filter out obs with no financial data
# dim(final_data)
```

```{r, message=FALSE, warning=FALSE}
final_data[cols] <- lapply(final_data[cols], as.numeric)
final_data$Bankrupt <- as.factor(final_data$Bankrupt)
final_data <- final_data %>%
  mutate(Bankrupt = ifelse(Bankrupt == 0, "No", "Yes"))
cols <- c("Closing_Price_Dec", "Closing_Price_Mar", "Closing_Price_Jun", "Closing_Price_Sep", "Market_Value", 
          "X3y.Price.Volatility", "X5y.Price.Volatility")
final_data[cols] <- lapply(final_data[cols], as.numeric)
```

```{r distribution}
final_data %>% 
  filter(!is.na(Total_Assets)) %>%
ggplot(aes(x = Total_Assets, fill = Bankrupt)) +
  geom_histogram(binwidth = 50, alpha=0.5, position="identity")

final_data %>%
  filter(!is.na(Net_Sales)) %>%
ggplot(aes(x = Net_Sales, fill = Bankrupt)) +
  geom_histogram(binwidth = 50, alpha=0.5, position="identity")

final_data %>%
  filter(!is.na(Market_Value)) %>%
ggplot(aes(x = Market_Value, fill = Bankrupt)) +
  geom_histogram(binwidth = 50, alpha=0.5, position="identity")
```

```{r revenue}
# in USD millions
# https://www.sangoma.com/articles/smb-sme-large-enterprise-size-business-matters/
final_data <- final_data %>%
  mutate(Size = case_when(
    Net_Sales <= 10 ~ "SMB", # 5-10 million
    Net_Sales <= 1000 ~ "SME", # 10 million - 1 billion
    Net_Sales > 1000 ~ "Large Enterprise") # over 1 billion
  )
final_data$Size <- as.factor(final_data$Size)
```

Companies with available revenue data fell under SMB or SME.

```{r}
final_data %>%
  group_by(Size, Bankrupt) %>%
  tally()
```

```{r mcap}
# in USD millions
# https://www.arborinvestmentplanner.com/market-capitalization-calculation-categories/
final_data <- final_data %>%
  mutate(Cap = case_when(
    Market_Value <= 300 ~ "Micro-cap", # 50-300 million
    Market_Value <= 2000 ~ "Small-cap", # 300 million - 2 billion
    Market_Value <= 10000 ~ "Mid-cap", # 2-10 billion
    Market_Value <= 200000 ~ "Large-cap", # 10-200 billion
    Market_Value > 200000 ~ "Mega-cap") # over 200 billion
  )
final_data$Cap <- as.factor(final_data$Cap)
```

Companies with available MCAP data fell under micro- or small-cap.

```{r}
final_data %>%
  group_by(Cap, Bankrupt) %>%
  tally()
```

```{r proportion}
final_data %>% 
  filter(!is.na(Size)) %>%
ggplot(aes(x = Size, fill = Bankrupt)) +
  geom_bar(position = "fill") +
  labs(x = "Category by Revenue Size", y = "Proportion", fill = "Bankrupt?")

final_data %>%
  filter(!is.na(Cap)) %>%
ggplot(aes(x = Cap, fill = Bankrupt)) +
  geom_bar(position = "fill") +
  labs(x = "Category by Market Cap", y = "Proportion", fill = "Bankrupt?")
```
Slightly higher proportion of companies are considered SME by revenue size and micro-cap by market capitalization. We will use revenue size (also highly correlated with total assets) and size of market capitalization as metrics for determining peer companies of bankrupt ones.

```{r}
# library("GGally")
# ggpairs(final_data, columns = c(9:10, 26), ggplot2::aes(colour = final_data$Bankrupt, alpha = 0.5))
# ggcorr(final_data, method = c("pairwise", "pearson"))
```

```{r}
final_data %>% 
  filter(!is.na(Total_Assets)) %>%
ggplot(aes(x = Total_Assets)) +
  facet_grid(Bankrupt~., scales = "free") +
  geom_histogram(binwidth = 50, alpha=0.5, position="identity")

final_data %>% 
  filter(!is.na(Net_Sales)) %>%
ggplot(aes(x = Net_Sales)) +
  facet_grid(Bankrupt~., scales = "free") +
  geom_histogram(binwidth = 50, alpha=0.5, position="identity")

final_data %>% 
  filter(!is.na(Market_Value)) %>%
ggplot(aes(x = Market_Value)) +
  facet_grid(Bankrupt~., scales = "free") +
  geom_histogram(binwidth = 50, alpha=0.5, position="identity")
```

Similar distribution of total asses, revenue, and MCAP for bankrupt and non-bankrupt companies.

```{r}
# final_data %>% 
#   filter(!is.na(Market_Value)) %>%
# ggplot(aes(x = Market_Value, fill = Bankrupt)) +
#   facet_wrap(.~Cap) +
#   geom_histogram(binwidth = 50, alpha=0.5, position="identity")
# 
# 
# final_data %>%
#   filter(!is.na(Total_Assets)) %>%
# ggplot(aes(x = Total_Assets, fill = Bankrupt)) +
#   facet_wrap(.~Size) +
#   geom_histogram(binwidth = 50, alpha=0.5, position="identity")
```

```{r}
# sector1 <- unique(final_data$FactSet.Econ.Sector)
# sector2 <- unique(final_data$Entity.NAICS.Sector.Full)
```

```{r}
# sector1 <- final_data %>%
#   group_by(Bankrupt, FactSet.Econ.Sector) %>%
#   tally()
```

```{r}
# sector1_name <- sector1[[2]][22:42]
# sector1_count <- sector1[[3]][22:42]
# sector1_prop <- sector1_count / 385
# list1 <- list(name = sector1_name,
#              count = sector1_count,
#              prop = sector1_count)
```

```{r}
final_data %>%
  group_by(Bankrupt) %>%
  tally()
```
385 Bankrupt
11944 Not Bankrupt

#### NAICS Sector
```{r 11 - Agriculture, Forestry, Fishing and Hunting}
sector <- "11 - Agriculture, Forestry, Fishing and Hunting"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB and Micro-cap (n=1)

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	and Micro-cap	(n=61)	
```

```{r 21 - Mining, Quarrying, and Oil and Gas Extraction}
sector <- "21 - Mining, Quarrying, and Oil and Gas Extraction"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	and Micro-cap	(n=22)		
# SME	and Micro-cap	(n=30)	
# SME	and Small-cap	(n=6)	

# SMB	NA	1		
# SME	NA	1		
# NA	Micro-cap	4		
# NA	Small-cap	1		
# NA	NA	2		

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB and	Micro-cap	(n=529)
# SME and	Micro-cap	(n=127)		
# SME	and Small-cap	(n=40)
```

```{r 22 - Utilities}
sector <- "22 - Utilities"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB and	Micro-cap	(n=1)		
# SME	and Micro-cap	(n=3)		

# NA	Micro-cap	1		
# NA	NA	1	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	and Micro-cap	(n=69)		
# SME	and Micro-cap	(n=23)		
```

```{r 23 - Construction}
sector <- "23 - Construction"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	and Micro-cap	(n=3)		
# SME	and Micro-cap	(n=3)

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	and Micro-cap	(n=73)		
# SME	and Micro-cap	(n=40)
```

```{r 31 - Manufacturing}
sector <- "31 - Manufacturing"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB and	Micro-cap	(n=55)		
# SMB	and Small-cap	(n=1)	
# SME	and Micro-cap	(n=56)		
# SME	and Small-cap	(n=7)		

# SMB	NA	5	
# SME	NA	2		
# NA	Micro-cap	5		
# NA	Small-cap	1		
# NA	NA	2	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	and Micro-cap	1333		
# SMB	and Small-cap	77		
# SME	and Micro-cap	823		
# SME	and Small-cap	317		
```

```{r 42 - Wholesale Trade}
sector <- "42 - Wholesale Trade"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	3		
# SME	Micro-cap	3		
# SME	Small-cap	1		

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	125		
# SME	Micro-cap	70	
# SME	Small-cap	4	
```

```{r 44 - Retail Trade}
sector <- "44 - Retail Trade"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	1		
# SME	Micro-cap	19		

# NA	Micro-cap	12		
# NA	Small-cap	4		
# NA	NA	3	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	106		
# SME	Micro-cap	72		
```
                                                         
```{r 48 - Transportation and Warehousing}
sector <- "48 - Transportation and Warehousing"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	1		
# SME	Micro-cap	6		

# NA	Micro-cap	1		
# NA	NA	1	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	26		
# SME	Micro-cap	73		
```

```{r 51 - Information}
sector <- "51 - Information"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	14		
# SME	Micro-cap	10	

# SMB	NA	1		
# NA	Micro-cap	1		
# NA	Small-cap	1

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	451		
# SME	Micro-cap	217		
```

```{r 52 - Finance and Insurance}
sector <- "52 - Finance and Insurance"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	7		
# SME	Micro-cap	13		
# SME	Small-cap	4

# SMB	NA	1
# SME	NA	1		
# NA	Micro-cap	1

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	604		
# SME	Micro-cap	758		
# SME	Small-cap	263	
```

```{r 53 - Real Estate and Rental and Leasing}
sector <- "53 - Real Estate and Rental and Leasing"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SME	Micro-cap	3		
# NA	Micro-cap	1	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SME	Micro-cap	60	
```
     
```{r 54 - Professional, Scientific, and Technical Services}
sector <- "54 - Professional, Scientific, and Technical Services"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	11		
# SME	Micro-cap	13		
# NA	Micro-cap	2	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	341		
# SME	Micro-cap	154		
```

```{r 55 - Management of Companies and Enterprises}
sector <- "55 - Management of Companies and Enterprises"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SME	Micro-cap	3	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SME	Micro-cap	183	
```

```{r 56 - Administrative and Support and Waste Management and Remediation Services}
sector <- "56 - Administrative and Support and Waste Management and Remediation Services"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	2		
# SME	Micro-cap	3		
# SME	Small-cap	1	
# SMB	NA	1		

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	130		
# SME	Micro-cap	60		
# SME	Small-cap	19		
```

```{r 61 - Educational Services}
sector <- "61 - Educational Services"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	1		
# SME	Micro-cap	1	

# SME	NA	1		
# NA	Micro-cap	1

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	18		
# SME	Micro-cap	16		
```
    
```{r 62 - Health Care and Social Assistance}
sector <- "62 - Health Care and Social Assistance"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	2		
# SME	Micro-cap	5
# SMB	NA	1	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	62		
# SME	Micro-cap	31		
```

```{r 71 - Arts, Entertainment, and Recreation}
sector <- "71 - Arts, Entertainment, and Recreation"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SME	Micro-cap	3		
# SME	NA	1	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SME	Micro-cap	23	
```
          
```{r 72 - Accommodation and Food Services}
sector <- "72 - Accommodation and Food Services"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	2		
# SME	Micro-cap	3	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	31		
# SME	Micro-cap	49		
```

```{r 81 - Other Services (except Public Administration)}
sector <- "81 - Other Services (except Public Administration)"
final_data %>%
  filter(Bankrupt == "Yes") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	1	

final_data %>%
  filter(Bankrupt == "No") %>%
  filter(Entity.NAICS.Sector.Full == sector) %>%
  group_by(Size, Cap) %>%
  tally()
# SMB	Micro-cap	29	
```
                                
```{r}
subset_final_data <- final_data %>%
  filter(Bankrupt == "Yes" |
          (Entity.NAICS.Sector.Full == "11 - Agriculture, Forestry, Fishing and Hunting" & 
         Size == "SMB" & Cap == "Micro-cap") |
          (Entity.NAICS.Sector.Full == "21 - Mining, Quarrying, and Oil and Gas Extraction" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap") | 
         (Size == "SME" & Cap == "Small-cap"))) |
          (Entity.NAICS.Sector.Full == "22 - Utilities" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "23 - Construction" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "31 - Manufacturing" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap") | 
         (Size == "SMB" & Cap == "Small-cap") | (Size == "SME" & Cap == "Small-cap"))) |
          (Entity.NAICS.Sector.Full == "42 - Wholesale Trade" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap") | 
         (Size == "SME" & Cap == "Small-cap"))) |
          (Entity.NAICS.Sector.Full == "44 - Retail Trade" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "48 - Transportation and Warehousing" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "51 - Information" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "52 - Finance and Insurance" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap") | 
         (Size == "SME" & Cap == "Small-cap"))) |
          (Entity.NAICS.Sector.Full == "53 - Real Estate and Rental and Leasing" & 
         Size == "SME" & Cap == "Micro-cap") |
          (Entity.NAICS.Sector.Full == "54 - Professional, Scientific, and Technical Services" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "55 - Management of Companies and Enterprises" & 
         Size == "SME" & Cap == "Micro-cap") |
          (Entity.NAICS.Sector.Full == "56 - Administrative and Support and Waste Management and Remediation Services" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap") | 
         (Size == "SME" & Cap == "Small-cap"))) |
          (Entity.NAICS.Sector.Full == "61 - Educational Services" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "62 - Health Care and Social Assistance" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "71 - Arts, Entertainment, and Recreation" & 
         Size == "SME" & Cap == "Micro-cap") |
          (Entity.NAICS.Sector.Full == "72 - Accommodation and Food Services" & 
         ((Size == "SMB" & Cap == "Micro-cap") | (Size == "SME" & Cap == "Micro-cap"))) |
          (Entity.NAICS.Sector.Full == "81 - Other Services (except Public Administration)" & 
         Size == "SMB" & Cap == "Micro-cap")
        )
```

```{r NAICS_sectors}
groups <- subset_final_data %>%
  group_by(Bankrupt, Entity.NAICS.Sector.Full) %>%
  tally()

subset_final_data %>%
  group_by(Bankrupt) %>%
  tally()
```

385 Bankrupt
7487 Not Bankrupt

```{r}
# Bankrupt = No
# groups1_name <- groups[[2]][1:19]
# groups1_count <- groups[[3]][1:19]
# groups1_prop <- groups1_count / 7487
# list1 <- list(name = groups1_name,
#              count = groups1_count,
#              prop = groups1_prop)
# 
# Bankrupt = Yes
# groups2_name <- groups[[2]][20:38]
# groups2_count <- groups[[3]][20:38]
# groups2_prop <- groups2_count / 385
# list2 <- list(name = groups2_name,
#              count = groups2_count,
#              prop = groups2_prop)
```

```{r}
# list <- list(name = groups1_name,
#              diff = groups2_prop - groups1_prop)
# 
# list
```
<!-- 52 - Finance and Insurance -->

```{r}
subset_final_data <- subset_final_data %>%
  mutate(Working_Capital_to_Total_Assets = 
           Working_Capital / Total_Assets,
         Retained_Earnings_to_Total_Assets = 
           Retained_Earnings / Total_Assets,
         EBITDA_to_Total_Assets = 
           EBITDA / Total_Assets,
         Sales_to_Total_Assets =
           Net_Sales / Total_Assets,
         EBITDA_Interest_Coverage_Ratio =
           EBITDA / Interest_Expense,
         Cash_Flow_to_Total_Debt = 
           Cash_From_Operations / Total_Debt,
         Market_Value_to_Total_Debt = 
           Market_Value / Total_Debt
         )
  
cleaned_data <- subset_final_data %>%
  select(1:3, 6:9, "ROA", "ROE", "Working_Capital_to_Total_Assets",
         "Retained_Earnings_to_Total_Assets", "EBITDA_to_Total_Assets",
         "Debt_to_EBITDA", "Sales_to_Total_Assets", "Profit_Margin",
         "Debt_to_Assets", "Debt_to_Equity", "Current_Ratio", "Interest_Coverage_Ratio",
         "EBITDA_Interest_Coverage_Ratio", "Cash_Flow_to_Total_Debt", "Market_Value_to_Total_Debt",
         27:33
         )
# dim(cleaned_data)
```

```{r match4}
# match market/macro data to the corresponding month and year
match4 <- function(year, month, list) {
  # data$Latest_Year, data$Month, e.g. cpi$CPILFESL
  matrix <- matrix(unlist(list), ncol = 12, byrow = TRUE) # row by year and col by month
  output <- rep(0, length(year))
  for (i in 1:length(year)) {
    row <- year[i] - 2009 # year (2010-2020)
    col <- month[i] # month (1-12)
    output[i] <- matrix[row, col]
  }
  return(output)
}
```

```{r}
list_cpi <- cpi$CPILFESL 
list_cpi <- list_cpi[-133] # remove 2021 data
list_yield <- yield$DGS3MO
list_yield <- list_yield[-c(133:134)] # remove 2021 data
list_oil <- oil$POILBREUSDM
list_oil <- list_oil[-133] # remove 2021 data
list_payroll <- nonfarm_payroll$Nonfarm.employment
list_payroll <- list_payroll[-133] # remove 2021 data
cleaned_data <- cleaned_data %>%
  mutate(CPI = match4(Latest_Year, Month, list_cpi),
         Yield = match4(Latest_Year, Month, list_yield),
         Oil = match4(Latest_Year, Month, list_oil),
         Nonfarm_Payroll = match4(Latest_Year, Month, list_payroll) # monthly changes in thousands
  )
```

```{r export_data}
# write.csv(cleaned_data, file = "cleaned_data.csv")
```